---
title: "Time Series Analysis: Utah"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Including Libraries
```{r}
library(xts)
library(tsbox)
```

# Reading the dataset
```{r}
weather <-read.csv("/Users/riya/Desktop/3144152.csv", as.is=T)
head(weather)
```
# Preprocessing the data

```{r}
weather[['DATE']] <- as.POSIXct(weather[['DATE']],format = "%Y-%m-%d")
```


```{r}
total_snow <- sum(weather$SNOW, na.rm=TRUE)
total_fog <- sum(weather$WT01, na.rm=TRUE)
total_thunder <- sum(weather$WT03, na.rm=TRUE)
total_icepellets <- sum(weather$WT04, na.rm=TRUE)
total_hail <- sum(weather$WT05, na.rm=TRUE)
total_glaze <- sum(weather$WT06, na.rm=TRUE)
total_damaging_winds <- sum(weather$WT11, na.rm=TRUE)
```

```{r}
sum_weather= total_snow+ total_fog +total_thunder+total_icepellets+ total_hail+ total_glaze+ total_damaging_winds
sum_weather
```
```{r}
per_snow=(total_snow/ sum_weather)*100
per_fog= (total_fog / sum_weather)*100
per_thunder= (total_thunder / sum_weather)*100
per_ice= (total_icepellets / sum_weather)*100
per_hail= (total_hail / sum_weather)*100
per_glaze= (total_glaze/ sum_weather)*100
per_winds= (total_damaging_winds / sum_weather)*100

```
```{r}
barplot(c(per_snow, per_fog, per_thunder, per_ice, per_hail, per_glaze, per_winds),
main = "weather conditions",
xlab = "conditions",
ylab = "percent",
names.arg = c("snow", "fog", "thunder", "icepellets", "hail", "glaze", "strong winds"),
col = "red",
horiz = FALSE)
```

```{r}
range(weather$DATE) 
```

# Preprocsessing the time series
#1.Creating time series ts2 of TMAX, TMIN, PRCP, SNOW
#2.Filling Null Values
#3.Start and end date window
```{r}
ts2 = xts(weather[,c("TMAX","TMIN","PRCP", "SNOW")], order.by=as.Date(weather$DATE))
#weather_ts
```
```{r}
ts2 = ts_xts(ts2)
#ts2
```
```{r}
ts2 = na.fill(ts2, "extend")

ts2 = window(ts2, start=as.Date("1995-01-01"), end=as.Date("2005-12-31"))

```

```{r}
ts_plot(ts2$SNOW, title="Snowfall")
```
```{r}
ts_plot(ts2$TMAX)
```




```{r}
plot(xts(ts2$TMAX),col="darkred", bty="n", las=1, fg=NA, ylim=c(-20, 120), ylab="Temperature (F)")

grid(nx=NA, ny=NULL, lty=1, col="gray")

legend("topright", fill=c("darkred", "navy"), cex=0.7,legend=c("TMAX", "TMIN"), bg="white")

```

```{r}
barplot(ts2$PRCP, border=NA, col="darkgreen", ylim=c(0, 2),space=0, bty="n", las=1, fg=NA, ylab="Daily Rainfall (inches)")

grid(nx=NA, ny=NULL, lty=1)

```

#Descriptive Stats

```{r}
summary(ts2)
```
# Reading with Minumum temperature
```{r}
ts2[ts2$TMIN == min(ts2$TMIN)]
```
# Reading with maximum temperature
```{r}
ts2[ts2$TMAX == max(ts2$TMAX)]
```

# Reading with maximum precipitation
```{r}
ts2[ts2$PRCP == max(ts2$PRCP)]
```
#seasonal Stats
```{r}
#Statistical summary of maximum temperature by months
ts2$MONTH = format(index(ts2), "%m")

months = split(as.numeric(ts2$TMAX), ts2$MONTH)

sapply(months, summary)
```
```{r}
#Statistical summary of minimum temperature by months
months = split(as.numeric(ts2$TMIN), ts2$MONTH)

sapply(months, summary)
```
```{r}
#Normalizing and Decomposing Precipitation
dat<- cumsum(rnorm(ts2$PRCP))
x<-ts(dat)
#stl(x, "periodic")
xx<- ts(dat, frequency=12)
decomp <-stl(xx, "periodic")
```
```{r}
plot(decomp)
```
```{r}
summary(decomp$time.series[,"trend"])
```

```{r}
#Normalizing and decomposing temperature
data_TMAX<- cumsum(rnorm(ts2$TMAX))
xx_TMAX<- ts(data_TMAX, frequency=12)
decomp_TMAX <-stl(xx_TMAX, "periodic")
plot(decomp_TMAX)
summary(decomp_TMAX$time.series[,"trend"])


```




#Aggregation
```{r}
#Aggregating Temperature
monthly.tmax = period.apply(ts2$TMAX, INDEX = seq(1, nrow(ts2) - 1, 30.4375), FUN = mean)

```

```{r}
data_tmax <-cumsum(rnorm(monthly.tmax))
xx_tmax<-ts(monthly.tmax, frequency=12)
```
```{r}
plot(ts_ts(xx_tmax),  col="darkred", ylim=c(20, 100), lwd=3, bty="n", las=1, fg=NA, ylab="TMAX (F)", main="Historical mean of Maximum temperatures")

grid(nx=NA, ny=NULL, lty=1)
```

```{r}


monthly.pmax = period.apply(ts2$PRCP, INDEX = seq(1, nrow(ts2) - 1, 30.4375), FUN = mean)

```


#Time series with frequency=12
```{r}

temperature.ts = ts(ts2$TMAX, start = c(1995, 01), end = c(2005, 12), frequency = 12)
plot(temperature.ts, xlab = "Year", ylab="temperature", main="temperature time series")
```

#Autocorrelation

```{r}
temperature.ts
lag.plot(temperature.ts, lags=12, do.lines=FALSE)
```



```{r}
lag.plot(temperature.ts, lag = 12, do.lines = FALSE)
acf(temperature.ts)
```
```{r}
#Partial Autocorrelation Function
pacf (temperature.ts, lag = 12 - 1, pl = TRUE)
```


```{r}
#Dickey Fuller test
# if p value > 0.05, the time series is not stationary
# if p value < 0.05, time series is stationary
#x <- rnorm(ts2)
#adf.test(temperature.ts, alternative = "stationary")
```












